---
title: "Lab 4 DDMII"
author: "Christine Brown"
date: "February 15, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  warning = F , message = F)
```

```{r}
setwd("C:/Users/brown/Documents/Grad School/Fall 16/DDMII/lab4")
library( maptools )
library( sp )
library( devtools )
library( dplyr )
library( DT )
library( pander )
syr <- readShapePoly( fn="01-05-2015", proj4string=CRS("+proj=longlat +datum=WGS84") )
syr.df <- as.data.frame( syr )
```


###Question 1: How many single family homes are in each neighborhood?

*Create a table of the count of single family homes by neighborhood*
```{r}
table( syr$Nhood[ syr$LandUse == "Single Family" ] )
```

*Highlight all single family homes, using a different color for each neighborhood*
```{r}
levels( syr.df$Nhood ) <- c( levels( syr.df$Nhood ),"Not Single Family" )
syr.df$Nhood[ syr.df$LandUse!="Single Family" ] <- "Not Single Family"
col.Nhood <- c( "lightpink", "hotpink", "deeppink", "violetred3", "indianred", "firebrick",
                "darkslategray2", "lightblue3", "dodgerblue", "royalblue2", "deepskyblue4", 
                "lightsalmon1", "darkorange", "chocolate1", "orange3", "orangered2",
                "mediumpurple1", "darkorchid1", "mediumorchid2", "magenta3", "darkmagenta",
                "yellow1", "lightgoldenrod1", "gold", "goldenrod", "darkgoldenrod1",
                "chartreuse2", "darkolivegreen3", "darkseagreen3", "green4", "seagreen4",
                "gray40", "gray94" )
palette( col.Nhood )
par(mar=c(0, 0, 2, 7) + 0.1)
plot( syr, col=syr.df$Nhood, border=F, main="Single Family Homes in Syracuse, NY" )
legend( 'bottomright', legend=levels(syr.df$Nhood), col=palette(),
        pch=19, pt.cex=.8, cex=0.6, xpd=NA, inset=c(-0.24,0))
```

###Question 2: Where does land in Syracuse have the highest value? 

*Create a table of the count of single family homes with values above $200k in each neighborhood, as a pecentage of all single family homes*
```{r}
SF <- syr.df[ syr.df$LandUse == "Single Family", ]
SF <- mutate( SF, Val200K = ifelse( AssessedVa>200000, "%>$200k", "%<$200k" ) )
round( prop.table( table( SF$Nhood, SF$Val200K ), 1 ) * 100, 2 ) %>% pander()
```

*Plot the value / acre of all parcels in Syracuse*
```{r}
syr.df <- mutate( syr.df, ValPerAcre = AssessedVa/Acres )
color.function <- colorRampPalette( c("firebrick4","gray88", "seagreen4" ) )
col.ramp <- color.function( 5 )
color.valperacre <- cut( rank(syr.df$ValPerAcre), breaks=5, labels=col.ramp )
color.valperacre <- as.character( color.valperacre )
par(mar=c(0, 0, 2, 7) + 0.1)
plot(syr, col=color.valperacre, border = F, main="Value per Acre of Land in Syracuse, NY")
breaks <- quantile(syr.df$ValPerAcre,probs = c(0, .2, .4, .6, .8, 1))
legend('bottomright', legend = c(paste("$",round(breaks[1],0),"- $",round(breaks[2],0)),
                                 paste("$",round(breaks[2],0),"- $",round(breaks[3],0)),
                                 paste("$",round(breaks[3],0),"- $",round(breaks[4],0)),
                                 paste("$",round(breaks[4],0),"- $",round(breaks[5],0)),
                                 paste("$",round(breaks[5],0),"- $",round(breaks[6],0))),
       col=col.ramp, pch=19, pt.cex=1.5, cex=0.7, xjust=1, xpd=NA, inset=c(-0.24,0))
```

###Question 3: What is the age of single family homes in each neighborhood? 

*Create a table that reports the 10th, 25th, 50th, 75th, and 90th percentile of home ages in each neighborhood.*
```{r}
SF$YearBuilt <- as.numeric(as.character(SF$YearBuilt))
SF <-  mutate(SF, Age = 2017-SF$YearBuilt)
SF_Grouped <- group_by(SF, Nhood)
summarise( SF_Grouped,
           "10th" = round( quantile( Age, probs = .1, na.rm=TRUE ), 0 ),
           "25th" = round( quantile( Age, probs = .25, na.rm=TRUE ), 0 ),
           "50th" = round( quantile( Age, probs = .5, na.rm=TRUE ), 0 ),
           "75th" = round( quantile( Age, probs = .75, na.rm=TRUE ), 0 ),
           "90th" = round( quantile( Age, probs = .9, na.rm=TRUE ), 0 )) %>% pander
```

*Create a choropleth map that shows the age of properties by decade, pre-1900s can be one category.*
```{r}
color.function.decade <- colorRampPalette( c( "lightblue1", "darkblue" ) )
col.ramp.decade <- color.function.decade( 13 )
color.decade <- cut( as.numeric( as.character( SF$YearBuilt ) ) , breaks = c( 1700 , seq( 1900, 2020, 10 ) ), labels=col.ramp.decade )
color.decade <- as.character( color.decade )
par( mar = c( 0, 0, 2, 7) + 0.1)
plot( syr[ syr$LandUse!="Single Family", ], border=FALSE, col="gray96",
      main="Construction Decade of Single Family Homes in Syracuse, NY", cex.main=1)
plot( syr[ syr$LandUse=="Single Family", ], border=FALSE, col=color.decade, add=TRUE)
legend( 'bottomright', legend = c("Pre-1900", "1900 - 1909", "1910 - 1919", "1920 - 1929",
                                   "1930 - 1939", "1940 - 1949", "1950 - 1959", "1960 - 1969", 
                                   "1970 - 1979", "1980 - 1989", "1990 - 1999", "2000 - 2009",
                                   "2010 - Present"),
       col=col.ramp.decade, pch=19, pt.cex=1.5, cex=0.75, xjust=1, xpd=NA, inset=c( -0.24,0 ) )